\Huge Handwritten digit recognition

### 2017-02-08 doc/Paper.tex
Talk about matrix multiplication

	\maketitle
	\setlength{\parindent}{0pt}

@@

\paragraph{Training} We employ gradient-descent on mini-batches. Early-stopping prevents the model from further learning when performance on the validation split starts decreasing.

@@

\paragraph{Evaluation}
TODO

@@

When looking at a batch of images:
\begin{itemize}
\item the batch is assigned evenly among workers,
\item each worker computes the gradient on its chunk of the batch,
\item partial gradients are sent from each worker to the master,
\item the master process accumulates all the partial gradients and aggregates them,
\item the master updates the model weights based on the total gradient,
\item updated weights are sent back to every worker,
\item a new batch is ready to be processed: each worker now knows the effects of all the others.
\end{itemize}

In our case, aggregating the partial gradients requires nothing more then a summation.
\\

@@

Another

TODO: talk about matrix multiplication overhead

@@

Matrices build as vector of vectors (using the standard library implementation), while providing greater flexibility, cannot be sent efficiently using MPI. MPI deals with contiguous blocks of memory, while STL vectors are, by definition, dynamic. A workaround this is to send a matrix row-by-row. Our initial implementation used this method but unfortunately it provides little to no increase in parallelization performance and caused a general-rewrite.

@@

The entire source code is available at \url{https://github.com/stefan1niculae/mnist-in-mpi}.

>>>>

	\maketitle
% 	\setlength{\parindent}{0pt}

@@

\paragraph{Training} We employ gradient-descent on mini-batches. Early-stopping prevents the model from further learning when performance on the validation split starts decreasing. Random sampling averts cycles when iterating through the dataset.

@@

\paragraph{Evaluation}
Model performance is measured on a separate testing set using the accuracy metric -- ie: the percentage of images correctly classified.

TODO

@@

When looking at a batch of images:
\begin{itemize}
\item the batch is assigned evenly among workers,
\item each worker computes the gradient on its chunk of the batch,
\item partial gradients are sent from each worker to the master,
\item the master process accumulates all the partial gradients and aggregates them,
\item the master updates the model weights based on the total gradient,
\item updated weights are sent back to every worker,
\item a new batch is ready to be processed: each worker now knows the effects of all the others.
\end{itemize}
In our case, aggregating the partial gradients requires nothing more then a summation.
\\

@@

Another approach would be to let the algorithm run sequentially and parallelize only the matrix multiplication operation. In light of the previously presented considerations, we can conclude that this would be grossly inefficient. The transfer overhead is much higher than the benefit gained by splitting the problem.

@@

Matrices build as vector of vectors (using the standard library implementation), while providing greater flexibility, cannot be sent efficiently using MPI. MPI deals with contiguous blocks of memory, while STL vectors are, by definition, dynamic. A work-around for this is to send a matrix row-by-row. Our initial implementation used this method but unfortunately it provides little to no increase in parallelization performance and caused a general-rewrite.

The most expensive matrix operations are dot products and exponentiation. For matrix multiplication, we explored the Strassen algorithm, $O(n^{2.8})$, and the Coppersmith–Winograd algorithm, $O(n^{2.37})$. Although they achieve high asymptotic performance, hidden constants make them impractical in our case. We went with the Naive, $\Theta(n^3)$, algorithm with the $ikj$ optimization.

@@

\noindent The entire source code is available at \url{https://github.com/stefan1niculae/mnist-in-mpi}.
